---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# geneticriskR

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/aballou16/geneticriskR.svg?branch=master)](https://travis-ci.org/aballou16/geneticriskR)
<!-- badges: end -->

The goal of geneticriskR is to provide standard formatting, statistical analyses
and plotting for polygenic risk score data generating using PRSice-2 software. 
Because recent genetic risk studies have shown a great deal of inconsistency, 
this package will allow for future studies to be more easily reproduced. 

## Installation

You can install the development version of geneticriskR from 
[Github](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("aballou16/geneticriskR")
```

## Example(s)

This is a basic example which demonstrates how to run a simple logistic 
regression model using an example output file from PRSice-2. 

```{r, eval = FALSE}
library(geneticriskR)
simple_mod <- simple_logistic_reg(test_data, hypertension)
simple_mod
```

This is another example which demonstrates how to run roc/auc analysis on the
model generated above. 
```{r, eval = FALSE}
library(geneticriskR)
proc_analysis(test_data, "hypertension", simple_mod)
```

